Rule-Based Preordering on Multiple Syntactic Levels \\in Statistical Machine Translation

### 2014-09-25 Paper/paper.tex
.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please make sure to keep technical paper submissions anonymous  !
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\name{Firstname Lastname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If multiple authors, uncomment and edit the lines shown below.       %%
%% Note that each line must be emphasized {\em } by itself.             %%
%% (by Stephen Martucci, author of spconf.sty).                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \makeatletter
 \def\name#1{\gdef\@name{#1\\}}
 \makeatother
 \name{{\em Ge Wu, Yuqi Zhang, Alex Waibel}}
%%%%%%%%%%%%%%% End of required multiple authors changes %%%%%%%%%%%%%%%%%

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technology, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu, alexander.waibel@kit.edu }
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the tree information of multiple syntactic levels. This approach extend the tree-based reordering from one level into multiple levels, which has the capability to process more complicated reordering cases. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no word reordering. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed further improvements of up to $0.43$ in BLEU score in the English-to-Chinese translation direction and further improvements of up to $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found many translation examples with improved syntactic structures.

@@

The rest of this paper is organized as follows. Section~\ref{re} presents a review of related work. In Section~\ref{mo}, we point out the problems for translation between English and Chinese and describe the motivation of MLT reordering. In Section~\ref{mu}, we introduce the details of the MLT preordering. In Section~\ref{ex}, we present the experiment results and evaluation. Finally, we conclude this paper in Section~\ref{co}.

@@

Reordering patterns are based on multiple levels of the syntax tree. Figure~\ref{unstructured} illustrates how the reordering patterns are detected from the syntax tree. In the example, the detection starts from the root node, go downwards three levels and use the nodes in these levels to detect the reordering pattern. These nodes that are used for detecting the reordering pattern are colored gray and have an italic font. Especially, the nodes with dark gray are the lowest ones among such nodes, which represent constituents that are actually reordered through the reordering rules. The leaf nodes in the syntax tree indicate words in the sentence, which has rectangle shape with angular corners.

@@

In order to evaluate the potential of word reordering, we also used oracle reordering in the experiments. Oracle reordering is considered to be an optimally reordered sentence as input to the SMT system and do not allow additional reordering during decoding \cite{combine}. The oracle reordering is created by using the permutation of source sentences, which is extracted from the word alignment between the source text and reference \cite{birch2}.

Because oracle reordering is considered as an optimally reordering, we can use it as input of the SMT system and the scores are the optimal results that can be achieved by word reordering, from which we can evaluate the potential of reordering methods.

@@

We have also found improvements in the sentence structure. Table~\ref{t1} and Table~\ref{t2} shows some translation examples in both translation directions. Sections are separated by double lines in the table. Each section of this table shows one translation example with the source sentence (\emph{source}), translation without using MLT reordering (\emph{no MLT}), translation with MLT reordering (\emph{MLT}) and the reference (\emph{reference}). \emph{Glossary} for the source sentences in translation examples from English to Chinese is also added for understanding. The translation without using MLT reordering comes from the configuration with highest BLEU score that didn't use MLT reordering. And the translation with MLT reordering comes from the configuration with highest BLEU score that used MLT reordering. From the examples, we can clearly see the improvements in sentence structure.

@@

\begin{table*}
\centering
\begin{tabular}{|l|m{0.7\textwidth}|} \hline
Source & \cntext{陈至立 说 , 古巴 是 拉美 和 加勒比 地区 有 重要 影响 的 国家 。}
\\ \hline
Glossary & chen-zhili said , cuba is latin-american and caribbean region has great influence of country .
\\ \hline
No MLT & chen zhili said : cuba is the latin america and the caribbean region has an important influence on the state .
\\ \hline
MLT & chen zhili said : cuba is a country of important influence latin america and the caribbean region .
\\ \hline
Reference & chen zhili said that cuba is a country of great influence in the latin american and caribbean region .
\\ \hline \hline

@@

\begin{table*}
\centering
\begin{tabular}{|l|m{0.7\textwidth}|} \hline
Source & hu jintao also extended deep condolences on the death of the chinese victims and expressed sincere sympathy to the bereaved families .
\\ \hline
No MLT & \cntext{胡锦涛 还 表示 深切 哀悼 的 受害者 家属 的 死亡 , 向 迂难者 家属 表示 诚挚 的 慰问 。} \\ \hline
MLT & \cntext{胡锦涛 还 对 中国 迂难者 表示 哀悼 , 向 迂难者 家属 表示 诚挚 的 慰问 。} \\ \hline 
Reference & \cntext{胡锦涛 还 对 中方 不幸 遇难 人员 表示 深切 的 哀悼 , 并 向 遇难 着 的 亲属 致以 诚挚 的 慰问 。} \\ \hline \hline
Source & satisfying personal interests and expanding knowledge are also major reasons why hourly work appeals to people .\\ \hline
No MLT & \cntext{满足 个人 利益 和 扩大 知识 也 是 主要 原因 小时 工作 吸引 人 。} \\ \hline
MLT & \cntext{满足 个人 利益 和 扩大 知识 也 是 为什么 学生 工作 吸引 人 的 主要 原因 。} \\ \hline 
Reference & \cntext{满足 个人 兴趣 , 扩大 自己 的 知识面 也 是 兼职 小时 工 受 青睐 的 一 个 重要 原因 。} \\ \hline \hline

Source & the dalai lama will go to visit washington this month .\\ \hline
No MLT & \cntext{达赖 喇嘛 将 访问 华盛顿 的 这 一 个 月 。} \\ \hline
MLT & \cntext{达赖 喇嘛 将 本 月 访问 华盛顿 。} \\ \hline
%? add more examples?
Reference & \cntext{达赖 喇嘛 将 在 本 月 前往 华盛顿 访问 。} \\ \hline
\end{tabular}
\caption{Examples of translations from English to Chinese}
\label{t2}
\end{table*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please make sure to keep technical paper submissions anonymous  !
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\name{Firstname Lastname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If multiple authors, uncomment and edit the lines shown below.       %%
%% Note that each line must be emphasized {\em } by itself.             %%
%% (by Stephen Martucci, author of spconf.sty).                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \makeatletter
 \def\name#1{\gdef\@name{#1\\}}
 \makeatother
% \name{{\em Ge Wu, Yuqi Zhang, Alex Waibel}}
%%%%%%%%%%%%%%% End of required multiple authors changes %%%%%%%%%%%%%%%%%

%\address{Institute for Anthropomatics \\
%Karlsruhe Institute of Technology, Germany \\
%{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu, alexander.waibel@kit.edu }
%}

\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the tree information of multiple syntactic levels. This approach extend the tree-based reordering from one level into multiple levels, which has the capability to process more complicated reordering cases. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no word reordering. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed further improvements of up to $0.43$ in BLEU score in the English-to-Chinese translation direction and further improvements of up to $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found many translation examples with improved syntactic structures.

@@

The rest of this paper is organized as follows. Section~\ref{re} presents a review of related work. In Section~\ref{mo}, we point out the problems for translation between English and Chinese and describe the motivation of MLT reordering. In Section~\ref{mu}, we introduce the details of the MLT reordering. In Section~\ref{ex}, we present the experimental results and evaluation. Finally, we conclude this paper in Section~\ref{co}.

@@

Reordering patterns are based on multiple levels of the syntax tree. Figure~\ref{unstructured} illustrates how the reordering patterns are detected from the syntax tree. In the example, the detection starts from the root node, go downwards three levels and use the nodes in these levels to detect the reordering pattern. These nodes that are used for detecting the reordering pattern are colored gray and have an italic font. The nodes with dark gray are the lowest ones among such nodes, which represent constituents that are actually reordered through the reordering rules. The leaf nodes in the syntax tree indicate words in the sentence, which has rectangle shape with angular corners.

@@

In order to evaluate the potential of word reordering, we also used oracle reordering in the experiments. Oracle reordering is considered to be an optimally reordered sentence as input to the SMT system and do not allow additional reordering during decoding \cite{combine}. So we can use it as input of the SMT system and the scores are the optimal results that can be achieved by word reordering, from which we can evaluate the potential of reordering methods.

@@

We have also found improvements in the sentence structure. Table~\ref{t1} and Table~\ref{t2} shows some translation examples in both translation directions. Sections are separated by double lines in the table. Each section of this table shows one translation example with the source sentence (\emph{source}), translation without using MLT reordering (\emph{no MLT}), translation with MLT reordering (\emph{MLT}) and the reference (\emph{reference}). \emph{Glossary} for the source sentences or references in Chinese is also added as word for word translation. Each word or group of hyphenated words in the glossary corresponds a Chinese character or a group of Chinese characters that are not separated with space. A placeholder $\Box$ is used to replace words that are difficult to translate, which play grammatical roles generally. The translation without using MLT reordering comes from the configuration with highest BLEU score that didn't use MLT reordering. And the translation with MLT reordering comes from the configuration with highest BLEU score that used MLT reordering. From the examples, we can clearly see the improvements in sentence structure.

@@

\begin{table*}
\centering
\begin{tabular}{|l|m{0.7\textwidth}|} \hline
Source & \cntext{陈至立 说 , 古巴 是 拉美 和 加勒比 地区 有 重要 影响 的 国家 。}
\\ \hline
Glossary & chen-zhili said , cuba is latin-american and caribbean region has great influence $\Box$ country .
\\ \hline
No MLT & chen zhili said : cuba is the latin america and the caribbean region has an important influence on the state .
\\ \hline
MLT & chen zhili said : cuba is a country of important influence latin america and the caribbean region .
\\ \hline
Reference & chen zhili said that cuba is a country of great influence in the latin american and caribbean region .
\\ \hline \hline

@@

\begin{table*}
\centering
\begin{tabular}{|l|m{0.7\textwidth}|} \hline
Source & hu jintao also extended deep condolences on the death of the chinese victims and expressed sincere sympathy to the bereaved families .
\\ \hline
No MLT & \cntext{胡锦涛 还 表示 深切 哀悼 的 受害者 家属 的 死亡 , 向 迂难者 家属 表示 诚挚 的 慰问 。} \\ \hline
MLT & \cntext{胡锦涛 还 对 中国 迂难者 表示 哀悼 , 向 迂难者 家属 表示 诚挚 的 慰问 。} \\ \hline 
Reference & \cntext{胡锦涛 还 对 中方 不幸 遇难 人员 表示 深切 的 哀悼 , 并 向 遇难者 的 亲属 致以 诚挚 的 慰问 。} \\ \hline
Glossary & hu-jintao also on chinese unfortunately killed people extended deep $\Box$ condolences , and to bereaved 's families expressed sincere $\Box$ sympathy .
\\ \hline \hline
Source & satisfying personal interests and expanding knowledge are also major reasons why hourly work appeals to people .\\ \hline
No MLT & \cntext{满足 个人 利益 和 扩大 知识 也 是 主要 原因 小时 工作 吸引 人 。} \\ \hline
MLT & \cntext{满足 个人 利益 和 扩大 知识 也 是 为什么 学生 工作 吸引 人 的 主要 原因 。} \\ \hline 
Reference & \cntext{满足 个人 兴趣 , 扩大 自己 的 知识面 也 是 兼职 小时 工 受 青睐 的 一 个 重要 原因 。}  \\ \hline
Glossary & satisfying personal interests , expanding own $\Box$ knowledge also are part-time hourly work is favored $\Box$ one $\Box$ major reason .
\\ \hline \hline

Source & the dalai lama will go to visit washington this month .\\ \hline
No MLT & \cntext{达赖 喇嘛 将 访问 华盛顿 的 这 一 个 月 。} \\ \hline
MLT & \cntext{达赖 喇嘛 将 本 月 访问 华盛顿 。} \\ \hline
%? add more examples?
Reference & \cntext{达赖 喇嘛 将 在 本 月 前往 华盛顿 访问 。} \\ \hline
Glossary & dalai lama will in this month go-to washington visit .
\\ \hline
\end{tabular}
\caption{Examples of translations from English to Chinese}
\label{t2}
\end{table*}
